### 2023-10-27 README.qmd
Add README.qmd, render in _targets, and tar_make()

---
format: gfm
---

# ManyEcoEvo <img src="man/figures/ManyEcoEvoHex.jpg" align="right" alt="" width="120" />

The `ManyEcoEvo` package provides a suite of functions for:

- Summarising, analysing and visualising the ManyEcoEvo dataset from 
*Gould et al.*[^1]
- Tidying and cleaning many-analyst style data for further analysis
- Reproducing the analysis in *Gould et al.* using your own many-analyst style data

> Note that the manuscript source-code for *Gould et al.* is located in a separate repository at [https://github.com/egouldo/ManyAnalysts](https://github.com/egouldo/ManyAnalysts), which can be viewed at [https://egouldo.github.io/ManyAnalysts/](https://egouldo.github.io/ManyAnalysts/).

## Installation

`ManyEcoEvo::` can be installed using [`devtools::`](https://devtools.r-lib.org) from GitHub with:

```{r eval=FALSE}
devtools::install_github("egouldo/ManyEcoEvo")
```

# Regenerating the ManyEcoEvo dataset

The data processing and analysis can be freely reproduced with the help of the [`targets::`](https://github.com/ropensci/targets) package. Please see the documentation at [https://docs.ropensci.org/targets/](https://docs.ropensci.org/targets/) for further detail.

Should you wish to completely reproduce the dataset generation and analysis in *Gould et al.*, complete the following steps:

1. Clone or download  [https://github.com/egouldo/this repository](https://github.com/egouldo/ManyEcoEvo)
2. Run `renv::restore()` to load the packages used in the analysis pipeline locally on your machine
3. Run `tar_destroy()` to remove any record and caches of existing targets
3. Run `targets::tar_make()` in your console, depending on the power of your machine, the analysis pipeline will take between 2 and 7 minutes to execute (plus or minus some!)
4. You can view a table of all targets in the pipeline by running `targets::tar_meta()`
5. To interact with objects or 'targets' within the analysis pipeline, call `targest::tar_load()` or `targets::tar_read()`:

```{r echo=FALSE}
suppressPackageStartupMessages(library(targets))
```

```{r}
targets::tar_read("ManyEcoEvo")
```

## License

This software is licensed with the GNU GPL 3 license [![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)

## Contributing

This package is released with a [Contributor Code of Conduct](https://github.com/egouldo/ManyEcoEvo/blob/aa2b9dcb6462f35ce873418e0b9c9697cf0b2f24/CODE_OF_CONDUCT.md) [![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-2.1-4baaaa.svg)](code_of_conduct.md)

## Citation

[^1]: Gould, E., Fraser, H., Parker, T. *et al.* (2023). Same data, different analysts: Variation in effect sizes due to analytical decisions in ecology and evolutionary biology [Preprint]. Ecology and Evolutionary Biology. https://doi.org/10.32942/X2GG62