### 2023-10-27 README.qmd
Add package citation info in README and render with tar_make()

---
format: gfm
---

# ManyEcoEvo <img src="man/figures/ManyEcoEvoHex.jpg" align="right" width="120"/>

The `ManyEcoEvo` package provides a suite of functions for:

-   Summarising, analysing and visualising the ManyEcoEvo dataset from *Gould et al.*[^1]
-   Tidying and cleaning many-analyst style data for further analysis
-   Reproducing the analysis in *Gould et al.* using your own many-analyst style data

> Note that the manuscript source-code for *Gould et al.* is located in a separate repository at <https://github.com/egouldo/ManyAnalysts>, which can be viewed at <https://egouldo.github.io/ManyAnalysts/>.

## Installation

`ManyEcoEvo::` can be installed using [`devtools::`](https://devtools.r-lib.org) from GitHub with:

```{r eval=FALSE}
devtools::install_github("egouldo/ManyEcoEvo")
```

# Regenerating the ManyEcoEvo dataset

The data processing and analysis can be freely reproduced with the help of the [`targets::`](https://github.com/ropensci/targets) package.
Please see the documentation at <https://docs.ropensci.org/targets/> for further detail.

Should you wish to completely reproduce the dataset generation and analysis in *Gould et al.*, complete the following steps:

1.  Clone or download [https://github.com/egouldo/this repository](https://github.com/egouldo/ManyEcoEvo)
2.  Run `renv::restore()` to load the packages used in the analysis pipeline locally on your machine
3.  Run `tar_destroy()` to remove any record and caches of existing targets
4.  Run `targets::tar_make()` in your console, depending on the power of your machine, the analysis pipeline will take between 2 and 7 minutes to execute (plus or minus some!)
5.  You can view a table of all targets in the pipeline by running `targets::tar_meta()`
6.  To interact with objects or 'targets' within the analysis pipeline, call `targest::tar_load()` or `targets::tar_read()`:

```{r echo=FALSE}
suppressPackageStartupMessages(library(targets))
```

```{r}
targets::tar_read("ManyEcoEvo")
```

## License

This software is licensed with the GNU GPL 3 license [![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)

## Contributing

This package is released with a [Contributor Code of Conduct](https://github.com/egouldo/ManyEcoEvo/blob/aa2b9dcb6462f35ce873418e0b9c9697cf0b2f24/CODE_OF_CONDUCT.md) [![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-2.1-4baaaa.svg)](code_of_conduct.md)

## Citation

```         
To cite package ‘ManyEcoEvo’ in publications use:

  Gould E, Fraser H, Nakagawa S, Parker T (2023). _ManyEcoEvo: Meta-analyse
  data from 'Many-Analysts' style studies_. R package version 1.0.0,
  <https://github.com/egouldo/ManyEcoEvo>.

A BibTeX entry for LaTeX users is

  @Manual{,
    title = {ManyEcoEvo: Meta-analyse data from 'Many-Analysts' style studies},
    author = {Elliot Gould and Hannah S. Fraser and Shinichi Nakagawa and Timothy H. Parker},
    year = {2023},
    note = {R package version 1.0.0},
    url = {https://github.com/egouldo/ManyEcoEvo},
  }
```

[^1]: Gould, E., Fraser, H., Parker, T.
    *et al.* (2023).
    Same data, different analysts: Variation in effect sizes due to analytical decisions in ecology and evolutionary biology \[Preprint\].
    Ecology and Evolutionary Biology.
    https://doi.org/10.32942/X2GG62