### 2023-11-10 manuscript/manuscript/manuscript.Rmd
Add text to summary section and statement of need section

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'ManyEcoEvo: an R package for working with the "Same Data, Different Analysts" Project in Ecology and Evolutionary Biology, and other many-analyst style data'
tags:
  - R
  - meta-analysis
  - metaresearch
  - ecology
  - evolutionary biology
  - many-analysts
authors:
  - name: Elliot Gould
    orcid: 0000-0003-0872-7098
    affiliation: "1" 
  - name: Hannah S. Fraser
    orcid: 0000-0000-0000-0000
    affiliation: "1"
  - name: Shinichi Nakagawa
    orcid: 0000-0000-0000-0000
    affiliation: "2"
  - name: Timothy H. Parker
    orcid: 0000-0000-0000-0000
    affiliation: "3"
affiliations:
 - name: The University of Melbourne
   index: 1
 - name: The University of New South Wales
   index: 2
 - name: Whitman College
   index: 3
citation_author: Gould and Fraser et al. 
date: "`r Sys.Date()`"
year: 2023
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

# Summary

Typical scientific approaches proceed with a single team, and one analytic specification, or at most several specifications implemented.
Thus the effect of alternative defensible analytic choices that could have been implemented is never known, either because alternative specifications aren't explored, or worse, because analysts / authors implement them but do not publish alternative analysis specifications.
However, we know from simulations and case studies that researcher degrees of freedom can result in spurious findings [@simmons2011], and these researcher degrees of freedom are susceptible to publication bias, p-hacking and other questionable research practice.
Similarly, specification curve analysis of published findings have demonstrated that the published conclusion had empirical support from only a small sample of the overall plausible specifications (evidence of spurious findings, and possibly publication bias and p-hacking) [@Simonsohn2019].

Many-analyst studies are a format of crowd-sourced scientific analysis [@schweinsberg2021], where analysts simultaneously, but independently, research the same research question, using the same dataset [@Voracek2019; @silberzahn2018].
Many-analyst studies demonstrate how subjective, often arbitrary and usually defensible analysis decisions may influence research results [@silberzahn2018] due to either disagreement among researchers or due to analytic uncertainty about choices such as which observations to include and what data-analysis strategy is most appropriate [@Voracek2019].

The *ManyEcoEvo* project is a many-analyst study that aimed to empirically explore the variation in effect sizes and model predictions generated by analytical decisions of different researchers within ecology and evolutionary biology [@gould2023].
174 analyst teams, comprising 246 analysts, independently investigated the answers to two pre-specified research questions using two previously unpublished datasets, one from conservation ecology, and one from evolutionary biology.
Analysts investigating the evolutionary biology dataset examined the effect of sibling number and nestling growth of blue tits (*Cyanistes caeruleus*), while analysts investigating the conservation ecology dataset examined the relationship between grass cover and *Eucalyptus* tree seedling recruitment.
The ManyEcoEvo project quantified the variation in results among analyses using meta-analysis and additionally sought to understand how features of studies such as, variable selection, random effects structures, qualitative and numeric ratings of the methods by peer-reviewers, influenced deviation from the meta-analytic mean.

*ManyEcoEvo* is an R package[@RCoreTeam2022] designed to *facilitate the analysis of* a many-analyst study in ecology and evolutionary biology, the *ManyEcoEvo Project* [@gould2023].
The purpose of *ManyEcoEvo* is to reproducibly implement the analysis and presentation of results for Gould *et al.* [-@gould2023] such that the 246 analysts as well as other interested researchers may independently reproduce the study's findings.
The package also aims to facilitate re-analyses and replications of Gould *et al.* [-@gould2023], as well as facilitating the analysis of similar many-analyst style datasets.

All raw data sets provided to independent teams for analysis, as well data collected on the analyses themselves are archived at **OSF Link1, OSF Link2,** respectively, and are provided in the GitHub repository for the package **(INSERT LINK)**, archived on Zenodo **(INSERT CITATION).** The ManyEcoEvo project contains tools for data-cleaning and processing tasks encountered when working with with many-analyst style data, functions for implementing and reproducing the analysis described inGould *et al.* [-@gould2023], establishes a structured workflow that facilitates re-analyses and replications of Gould *et al.* [-@gould2023] as well as implementation for new many-analyst studies, including a planned extension of the ManyEcoEvo project.

# Statement of Need

The *ManyEcoEvo* package was created to implement the analysis for the 'Same Data, Many Analysts in Ecology and Evolutionary Biology' project [@gould2023] and to allow the 246 co-authors and other interested researchers to computationally reproduce our analysis.
Following [@Marwick2018], [@Vuorre2021] and [@Filazzola2022], we opted to develop this research compendium as an R package as we find it best encapsulates the principles of 'clean code' for effective communication of the science, and therefore facilitates future reproduction, replication and potential synthesis [R packages as research compendia].
The results of some many-analyst studies have been controversial, yielding re-analyses criticising the original implementation of some many-analyst studies.
For example, [@Auspurg2021] re-analysed part of the seminal 'Red-Card' many-analyst study examining the influence of whether referees were more likely to give red cards to soccer players with dark skin tone than those with light skin tone [@silberzahn2018].
By providing the code and data for Gould *et al.* [-@gould2023] the *ManyEcoEvo* package, we aim to facilitate computational reproductions by others in order to verify our results.
We also aim to facilitate analytic reproductions of this study, wherein *ManyEcoEvo* can be used to re-run the analysis in Gould *et al.* [-@gould2023] with slightly different methods, or subsets of data in order to verify and examine the robustness of our findings.

# Same Data, Different Analysts: variation in effect sizes due to analytical decisions in ecology and evolutionary biology.

# Acknowledgements

We acknowledge the contributions of all 246 analysts and co-authors of Gould et al. (2023).
EG is supported by an Australian Government Research Training Program Scholarship.

-   [ ] HF?
-   [ ] SN?
-   [ ] TP?

# References